---
title: "Week1"
author: "JS"
date: "12/4/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Week 1 
We're going to use the R programming language to analyze social network data. The codes below is an adaptation of Irizarry & Love's e-book, http://genomicsclass.github.io/book/


## Installing R

The first step is to install R. You can download and install R from
the [Comprehensive R Archive Network](https://cran.r-project.org/)
(CRAN). It is relatively straightforward, but if you need further help
you can try the following resources: 

* [Installing R on Windows](https://github.com/genomicsclass/windows#installing-r) 
* [Installing R on Mac](http://youtu.be/Icawuhf0Yqo)
* [Installing R on Ubuntu](http://cran.r-project.org/bin/linux/ubuntu/README)


## Installing RStudio

The next step is to install RStudio, a program for viewing and running
R scripts. Technically you can run all the code shown here without
installing RStudio, but we highly recommend this integrated
development environment (IDE). Instructions are
[here](http://www.rstudio.com/products/rstudio/download/) and for
Windows we have special
[instructions](https://github.com/genomicsclass/windows). 

## Learn R Basics

Two key things you need to know about R is that you can get help for a function using `help` or `?`, like this:
```{r,eval=FALSE}
?install.packages
help("install.packages")
```

## Installing Packages

The first R command we will run is `install.packages`. R only includes a basic set of
functions. It can do much more than this, but not everybody needs
everything so we instead make some functions available via
packages. Many of these functions are stored in CRAN. Note that these
packages are vetted: they are checked for common errors and they must have
a dedicated maintainer. You can easily install packages from within R if you know
the name of the packages. 

```{r,eval=FALSE}
install.packages("dplyr")
```

We can then load the package into our R sessions using the `library` function:

```{r}
library(dplyr)
```

From now on you will see that we sometimes load packages without
installing them. This is because once you install the package, it
remains in place and only needs to be loaded with `library`. If you
try to load a package and get an error, it probably means you need to
install it first.

## Importing Data into R

The first step when preparing to analyze data is to read in the data into R. 

Although there are R packages designed to read Excel (xls) format, you generally want
to avoid this and save files as comma delimited (Comma-Separated
Value/CSV) or tab delimited (Tab-Separated Value/TXT) files.
These plain-text formats are often easier for sharing data with
collaborators, as commercial software is not required for viewing or
working with the data.

Also before you start thinking aobut how to load your files into R, you need to make sure that your data is well prepared to be imported. 1) Make sure your first row of your dataset is reserved for the header (variable names), while the frist column is used to identify the unit (ID). 2) Avoid names, values, or fields with blank spaces, otherwise each word will be interpreted as a spearate variable, resulting in errors. If you want to concatenate words, do this by inserting a "."" or "_"" 

We will start with a simple example dataset from Katya Ognyanova's website. This dataset involves a network of hyperlinks and mentions among news sources. 
"Ognyanova_Media_NODES.csv" and "Ognyanova_Media_EDGES.csv" You can download the datasets from here (https://github.com/jieunshi/sna_class). Save this file in a folder that you want to work with for this class. 

Now, the next step is to find the file containing your data and know its *path*. 


#### Paths and the Working Directory

When you are working in R it is useful to know your _working directory_. This is the directory or folder in which R will save or look for files by default. You can see your working directory by typing:

```{r, eval=false}
getwd()
```

You can also change your working directory using the function `setwd`. Or you can change it through RStudio by clicking on "Session". 

```{r, eval=false}
setwd("/Users/jieun.shin/Documents/Teaching/SNA/Week1_Intro")
```

The functions that read and write files (there are several in R) assume you mean to look for files or write files in the working directory. Our recommended approach for beginners will have you reading and writing to the working directory. 

### Reading your csv data file 
And now we are ready to read in the file. 
The "read.csv"" function will read your "csv" file. 

```{r,eval=FALSE}
dat <- read.csv("Ognyanova_Media_NODES.csv")
```

If you did not receive any message, then you probably read in the file successfully.

### Reading your excel data file
If you have an excel file, install the "xlsx" package and use the function "read.xlsx"

```{r,eval=FALSE}
install.packages("xlsx")
library(xlsx)
dat <- read.xlsx("put_your_file_name_here.xlsx")
```

### Descriptive statistics with dplyr 

Let's start with calculating mean, standard deviation, and median of audience sizes in the dataset 

```{r,eval=FALSE}
mean(dat$audience.size)
sd(dat$audience.size)
median(dat$audience.size)
```

The summarize function in dplyr provides a way to compute sumary statistics with intutitive and readable code. 

```{r,eval=FALSE}
s <- dat %>%
  summarize(average=mean(dat$audience.size), standard_deviation=sd(dat$audience.size), median =median(dat$audience.size))

s
```


The codes below using the pipe operator will acheive the same results 

```{r,eval=FALSE}
s <- dat %>%
  summarize(average=mean(audience.size), standard_deviation=sd(audience.size), median =median(audience.size))

s
```

Sometimes we want to first split data into groups and then compute summaries for each group. 
For example, we may want to compute the average and standard deviation for newspapers, TV, and online. For this we will use, the "group_by" operator. 


```{r,eval=FALSE}
q <- dat %>%
  group_by(type.label) %>%
  summarize(average=mean(audience.size), standard_deviation=sd(audience.size), median =median(audience.size))

q 
```

### Visualization with ggplot2 

you can do this with ggplot2. 

```{r,eval=FALSE}
library(ggplot2)

ggplot(dat, aes(x=type.label, y=audience.size)) +geom_point()

```

end 