---
title: "Week4_Rcode"
author: "JS"
date: "12/14/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## This session covers basics of network analysis and visualization with the igraph package.
## igraph is a collection of R functions to explore, create, manipulate and visualize graphs. 
## Let's first install the latest version of igraph 

```{r, eval=false}
install.packages("igraph")
library(igraph)
```

## We will use the datasets and tutorials created by Ognyanova. 
## Download the two files "Media_nodes.csv" and "Media_edges.csv" from github. 
## Don't forget to set your working directory. 
## You need to replace the following code with your own path to your file location 

setwd("/Users/jieun.shin/Documents/Teaching/SNA/Week4_centrality")

## "Media_nodes.csv" contains characteristics of each media organizations 
## "Media_edges.csv"" contains relationships among media organizations. 

```{r, eval=false}
nodes <-read.csv(file="Media_nodes.csv", header=TRUE, sep="\t")
links <-read.csv(file="Media_edges.csv", header=TRUE, sep="\t")
```

## Since "links"" datset is an edgelist, we need to convert it to a network object. 
## igraph has a convenient function called "graph_from_data_frame" to do this task 

```{r, eval=false}
net <- graph_from_data_frame(d=links, vertices=nodes, directed=T)
net
```


## Check basic network measures: size, density, centrliazation, etc 

```{r, eval=false}
vcount(net)
ecount(net)
graph.density(net)
centr_degree(net)
```

## Let's make our first attempt to plot it 
```{r, eval=false}
plot(net)
```

## The visualization shows self-loops (relationship with oneself). 
## We will remove the loops in the graph 
```{r, eval=false}
net <- simplify(net)
```

## See if it worked 
```{r, eval=false}
plot(net)
```

## The graph looks too crowded. We can do better. 

```{r, eval=false}
plot(net, edge.arrow.size=.4, vertex.label=NA)
```

## A few more adjustments 

```{r, eval=false}
plot(net, edge.arrow.size=.2, edge.color="orange",
vertex.color="orange", vertex.frame.color="#ffffff",
vertex.label=V(net)$media, vertex.label.color="black")
```

## Examine indegree centrality scores 
```{r, eval=false}
degree(net, mode="in")
```

## Plot network based on this indegree measure. 
##Let's make the node size proportional to indegree. We will also remove vertex lables 
```{r, eval=false}
V(net)$indegree <-degree(net, mode="in")
plot(net, vertex.size=V(net)$indegree)
```

## We need to make vertex sizes a bit bigger and arrows smaller 
```{r, eval=false}
plot(net, vertex.size=V(net)$indegree*5, vertex.label=NA, edge.arrow.size=.2)
```

## If you want to print the vertex's real name, use the code below. 
## Set each label's text size to be .7 
```{r, eval=false}
plot(net, vertex.size=V(net)$indegree*5, vertex.label=V(net)$media, vertex.label.cex=.7, edge.arrow.size=.2)
```

## To color the nodes based on their attributes, first you need to learn the logic 
## Let's color them based on media.type
```{r, eval=false}
V(net)$color <-"black"
V(net)$color <- ifelse(V(net)$media.type==1, "blue", V(net)$color)
V(net)$color <- ifelse(V(net)$media.type==2, "red", V(net)$color)
V(net)$color <- ifelse(V(net)$media.type==3, "orange", V(net)$color)
plot(net, vertex.size=V(net)$indegree*5, vertex.color=V(net)$color, vertex.label=V(net)$media, vertex.label.cex=.7, edge.arrow.size=.2)
```

## There are other ways to generate the colors 
```{r, eval=false}
colrs <-c("blue", "red", "orange")
V(net)$color <- colrs[V(net)$media.type]

plot(net, vertex.size=V(net)$indegree*5, vertex.color=V(net)$color, 
vertex.label=NA, vertex.label.cex=.7, edge.arrow.size=.2, edge.width=E(net)$width, main="Network of Media Organizations")
```

## Set edge width based edge weight. We will get rid of vertex labels 
```{r, eval=false}
V(net)$color <-"black"
V(net)$color <- ifelse(V(net)$media.type==1, "blue", V(net)$color)
V(net)$color <- ifelse(V(net)$media.type==2, "red", V(net)$color)
V(net)$color <- ifelse(V(net)$media.type==3, "orange", V(net)$color)
E(net)$width <- E(net)$weight/6

plot(net, vertex.size=V(net)$indegree*5, vertex.color=V(net)$color, 
vertex.label=NA, vertex.label.cex=.7, edge.arrow.size=.2, edge.width=E(net)$width)
```

## Add a title 
```{r, eval=false}
V(net)$color <-"black"
V(net)$color <- ifelse(V(net)$media.type==1, "blue", V(net)$color)
V(net)$color <- ifelse(V(net)$media.type==2, "red", V(net)$color)
V(net)$color <- ifelse(V(net)$media.type==3, "orange", V(net)$color)
E(net)$width <- E(net)$weight/6

plot(net, vertex.size=V(net)$indegree*5, vertex.color=V(net)$color, 
vertex.label=NA, vertex.label.cex=.7, edge.arrow.size=.2, edge.width=E(net)$width, main="Network of Media Organizations")
```

## Add a legend
```{r, eval=false}
colrs <-c("blue", "red", "orange")
V(net)$color <- colrs[V(net)$media.type]

plot(net, vertex.size=V(net)$indegree*5, vertex.color=V(net)$color, 
vertex.label=NA, vertex.label.cex=.7, edge.arrow.size=.2, edge.width=E(net)$width, main="Network of Media Organizations")
legend("bottomright", c("Newspaper","Television","Online News"), col=colrs, pch=16)
```
