---
title: "week3_rcode"
author: "JS"
date: "12/12/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Let's install package 'egor'
## This R package simplifies many common ego-network analysis tasks
## More information can be found here https://cran.r-project.org/web/packages/egor/egor.pdf

```{r, eval=false}
install.packages("egor")
library(egor)
```

# An "egor" object contains all data levels associated with ego-centered network analysis such as (1)ego, (2)alter, (3) alter-alter ties. By providing the "egor()" function, we construct an egor object. The data should contain the three types of data levels. We can take a look at the examples provided by the egor package. Pay attention to the ID variables connecting the levels with each other. 


```{r, eval=false}
data("alters32")
data("egos32")
data("edges32")
```

#Let's combine the data so that we can create an egor object. 

```{r, eval=false}
e1 <-egor(alters.df=alters32, egos.df=egos32, aaties=edges32, 
          ID.vars=list(ego="egoID", alter ="alterID", source="Source", target="Target"))

e1
```

#use the following functions to analyse the networks 

```{r, eval=false}
summary(e1)
```


```{r, eval=false}
ego_density(e1)
```

# diversity 

```{r, eval=false}
alts_diversity_entropy(e1, "alter.age")
```

#Ego-Alter Homophily (EI-Index)

```{r, eval=false}
comp_ei(e1, "alter.age", "age")
```


## visualization 
```{r, eval=false}
data(egor32)
egor_vis_app(egor32)
```


